Signatures of Foreground Sky in Power Spectrum of Redshifted Neutral Hydrogen from the Epoch of Reionization

### 2014-09-20 foreground_delay_spectrum_modeling.tex
From v0.6: =========
(*) Made minor reshuffling of paragraphs up to section 4.
(*) Minor edits up to section 4.

\def\UW{\altaffilmark{2}}
\def\UWtxt{\altaffiltext{2}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}

\def\RRI{\altaffilmark{3}}
\def\RRItxt{\altaffiltext{3}{Raman Research Institute, Bangalore 560080, India}}

\def\SKASA{\altaffilmark{4}}
\def\SKASAtxt{\altaffiltext{4}{Square Kilometre Array South Africa (SKA SA), Cape Town 7405, South Africa}}

\def\ANU{\altaffilmark{5}}
\def\ANUtxt{\altaffiltext{5}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}

\def\Haystack{\altaffilmark{6}}
\def\Haystacktxt{\altaffiltext{6}{MIT Haystack Observatory, Westford, MA 01886, USA}}

\def\Curtin{\altaffilmark{7}}
\def\Curtintxt{\altaffiltext{7}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}

\def\USydney{\altaffilmark{8}}
\def\USydneytxt{\altaffiltext{8}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia}}

\def\CAASTRO{\altaffilmark{9}}
\def\CAASTROtxt{\altaffiltext{9}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}

\def\MIT{\altaffilmark{10}}
\def\MITtxt{\altaffiltext{10}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}

\def\CfA{\altaffilmark{11}}
\def\CfAtxt{\altaffiltext{11}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}

@@

%% Author list
\author{
%% Lead Authors
Nithyanandan~Thyagarajan\ASU,
Daniel~C.~Jacobs\ASU,
Judd~D.~Bowman\ASU,
%% Supporting authors
A.~P.~Beardsley\UW,
B.~J.~Hazelton\UW, 
M.~F.~Morales\UW, 
J.~Pober\UW,
I.~S.~Sullivan\UW,
T.~Prabu\RRI, 
%% Builders' list
G.~Bernardi\SKASA,
F.~Briggs\ANU,
R.~J.~Cappallo\Haystack, 
B.~E.~Corey\Haystack, 
A.~A.~Deshpande\RRI, 
D.~Emrich\Curtin,
B.~M.~Gaensler\USydney$^,$\CAASTRO, 
R.~Goeke\MIT,
L.~J.~Greenhill\CfA,
M.~Johnston-Hollitt\Victoria,
D.~L.~Kaplan\UWisc, 
J.~C.~Kasper\UMichigan$^,$\CfA, 
E.~Kratzenberg\Haystack, 
C.~J.~Lonsdale\Haystack, 
M.~J.~Lynch\Curtin, 
S.~R.~McWhirter\Haystack,
D.~A.~Mitchell\CASS$^,$\CAASTRO, 
E.~Morgan\MIT, 
D.~Oberoi\Tata, 
S.~M.~Ord\Curtin$^,$\CAASTRO,
A.~E.~E.~Rogers\Haystack, 
A.~Roshi\NRAO, 
N.~Udaya~Shankar\RRI, 
K.~S.~Srivani\RRI, 
R.~Subrahmanyan\RRI$^,$\CAASTRO, 
S.~J.~Tingay\Curtin$^,$\CAASTRO, 
M.~Waterson\Curtin$^,$\ANU,
R.~B.~Wayth\Curtin$^,$\CAASTRO, 
R.~L.~Webster\UMelbourne$^,$\CAASTRO, 
A.~R.~Whitney\Haystack, 
A.~Williams\Curtin, 
C.~L.~Williams\MIT
}

@@

Detection of 21~cm H{\sc i} emission from the Epoch of Reionization (EoR), at redshifts $z>6$, is limited primarily by foreground emission.  We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.  Using the delay spectrum technique, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal via spillover outside the {\it wedge} arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array (MWA) observations with simulations separated into components based on sky position and type of emission, we identify diffuse emission near the horizon as the largest contributing factor, even on wide antenna spacings.   For signals entering through the primary field of view, compact objects dominate the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the {\it wedge} in Fourier space.  Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate a large fraction of foreground contamination in redshifted 21~cm power spectrum analysis.

@@

After the CMB epoch, the period in the history of the universe often referred to as the {\it dark ages}, followed the epoch of reionization (EoR). This was a period of non-linear growth of density perturbations and astrophysical evolution. Studying the epoch of reionization holds the key to understanding this evolution. Recently, observing the redshifted 21~cm spin transition of neutral hydrogen has emerged as a very promising experiment to fill the gaps in our understanding of the universe's history.

@@

Perhaps the key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}. Considerable effort has been made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,liu11,dat10,bow09,liu09}. A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes lost by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components most likely to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.

Foregrounds contaminate the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analysis. Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity. To decouple our findings from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline-based approach of \citet{pob13,par14}. This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.

Matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline. We find that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon. We find that diffuse Galactic emission plays a larger role then previously thought. We quantify this contamination by separating the simulations into components based on type of foreground emission. Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

@@

Interferometer array data known as {\it visibilities}, $V_f(\overline{\mathbf{u}},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\overline{\mathbf{x}}$ and then Fourier transformed along time axis to obtain a spectrum along frequency ($f$) axis. Here, $\overline{\mathbf{u}}\equiv \overline{\mathbf{x}}f/c \equiv \overline{\mathbf{x}}/\lambda$, $c$ is the speed of light and $\lambda$ is the wavelength. $\overline{\mathbf{u}}$ is related to transverse spatial frequency modes ($\overline{\mathbf{k}}_\perp$) of brightness distribution of the sky as $\overline{\mathbf{k}}_\perp \equiv 2\pi\overline{\mathbf{u}}/D(z)$, where $D(z)$ is the transverse comoving distance at a redshift $z$. If $I(\overline{\mathbf{l}},f)$ is the emission at different frequencies on the sky as a function of direction unit vector ($\overline{\mathbf{l}}$) specified as direction cosines, $V_f(\overline{\mathbf{u}},f)$ represents the Fourier decomposition of $I(\overline{\mathbf{l}},f)$ attenuated by the antenna's angular power--pattern $A(\overline{\mathbf{l}},f)$, into transverse spatial frequency modes. Taking into account the instrumental bandpass weights $W_f(f)$, this relation can be written as:
\begin{align}\label{eqn:obsvis}
  V_f(\overline{\mathbf{u}},f) &= \int A(\overline{\mathbf{l}},f)\,I(\overline{\mathbf{l}},f)\,W_f(f)\,e^{-i2\pi \overline{\mathbf{u}}\cdot\overline{\mathbf{l}}}\,\dif^2 \overline{\mathbf{l}}.
\end{align}

We define the {\it delay spectrum} $V_\eta(\overline{\mathbf{u}},\eta)$ to be the inverse Fourier transform of $V_f(\overline{\mathbf{u}},f)$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  V_\eta(\overline{\mathbf{u}},\eta) &= \int V_f(\overline{\mathbf{u}},f)\,W'_f(f)\,e^{i2\pi\eta f}\,\dif f,
\end{align}
where, $W'_f(f)$ is a frequency window weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}. $\eta$ represents the signal delay between the antenna pairs in an interferometer given by:
\begin{equation}\label{eqn:delay}
  \eta = \frac{\overline{\mathbf{x}}\cdot\overline{\mathbf{l}}}{c} = \frac{\overline{\mathbf{u}}\cdot\overline{\mathbf{l}}}{f}.
\end{equation}
Note that as the visibility is a fully complex quantity, its Fourier transform is fully Nyquist sampled over positive and negative delays. Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna pair's baseline vector. For example, a perfectly Eastward oriented baseline vector will observe an object in the Eastern sky at positive delay, and conversely for an object in the Western sky.

Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight. Consequently, $\eta$ is a measure of line of sight spatial frequency mode, $k_\parallel$, given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. We use $H_0=100\,h$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.

Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of redshifted H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ \citep{mor04}. However, foreground objects emit in radio frequencies and contaminate the signals from the redshifted H{\sc i}. Due to their spatial and spectral properties, $V_\eta(\overline{\mathbf{u}},\eta)$ also contains contamination from foreground radio objects \citep{thy13,tro12,mor12,bow09}. These contaminations have to be characterized precisely in order to reduce their impact on redshifted H{\sc i} power spectrum detection sensitivity.

In summary, $V_\eta(\overline{\mathbf{u}},\eta)$, is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by each antenna pair, as seen from equation~\ref{eqn:delay-transform}. It captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is very closely related to the sought power $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ in $k$--modes containing critical information about spatial scales. For our analysis, we use an equivalent quantity, $V_\eta(\overline{\mathbf{x}},\eta) \equiv V_\eta(\lambda\overline{\mathbf{u}},\eta)$. For purposes of illustration, we often use a simplified quantity $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$.

Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper. $|\overline{\mathbf{x}}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis. $\eta$ and $k_\parallel$, denoting spatial scales along line of sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge}, the boundaries of which are determined by antenna spacing and the light travel time across it. These boundaries are called horizon delay limits and are shown by solid lines. Spectral response of the instrument convolves the foreground wedge and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\eta$--axis. This effect has been quantified in \citet{thy13}. The width of this narrow strip is inversely proportional to operating bandwidth. Region of Fourier space excluding the {\it foreground wedge} and also the narrow strips is usually called the {\it EoR window}. These are shown in light and medium shades. However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape that results in harmonic repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s. The underlying cause of this effect will be discussed later in this paper. Hence, we have identified a region of high significance (shown in medium shade) that forms a subset of the {\it EoR window} bounded by $|\eta|\le 0.78\,\mu$s where contamination from foregrounds is expected to be low and yet the EoR H{\sc i} power spectrum signal is reasonably strong. Hereafter, we refer to this region as the {\it EoR window} in the rest of this paper. The MWA EoR collaboration actively uses this as a test region in Fourier space for diagnostics on data quality.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
\caption{Illustration of Fourier space in which delay (and power) spectra of H{\sc i} signals from the EoR are presented. The $x$--axis denoted by $|\overline{\mathbf{x}}|$ (antenna spacing in m) or $k_\perp$ (transverse wavenumber in units of $h$~Mpc$^{-1}$) are equivalent and represent transverse spatial scales. Similarly, the $y$--axis denoted by $\eta$ (delay in s or $\mu$s) or equivalently, $k_\parallel$ (line--of--sight wavenumber in units of $h$~Mpc$^{-1}$) represents spatial scales along the line of sight. The dark shaded region is often referred to as {\it foreground wedge} where smooth spectrum foregrounds reside. The boundaries of the {\it foreground wedge} (solid lines) are given by the light travel time for corresponding antenna spacings and is referred to as the horizon delay limits. In principle, region outside the wedge is expected to be relatively free of foreground contamination and is called the {\it EoR window} (light and medium shades). In practice, this contamination is also high in the narrow unshaded white regions (bound by solid and dashed lines) due to convolution by instrumental transfer function. The width of this narrow region is determined by inverse of operating bandwidth. Due to an undesirable spectral response of the instrument used in this study, namely the MWA, we identify a narrower region of high significance to EoR studies (medium shade) and hereafter refer to it as the {\it EoR window} in this paper. \label{fig:fourier-space}}
\end{figure}

Throughout this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site. For the conventions of Fourier transform and its inverse we use in this paper, in accordance with equation~\ref{eqn:delay}, signals from the sky towards east and north are recorded with positive delays on eastward and northward oriented antenna spacings respectively. For instance, at the beginning of the observation, the Galactic center is in the westward sky just about to set. Signals from this direction are recorded with a negative delay. Similarly, signals from directions eastward of the local meridian arrive with positive delays.

For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all antenna spacings to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$. Antenna spacings oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.

@@

The MWA was completed in 2012, and after commissioning, began its EoR observing program in 2013. In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30~MHz band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles each in the shape of a bow--tie characterized by a grating response with a $\gtrsim$~20\arcdeg wide primary field of view and multiple secondary side lobes. The tile arrangement is a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius with the longest baselines at 3~km \citep{bea12}. Here we focus on antenna spacings $|\overline{\mathbf{x}}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz. The center frequency 185~MHz corresponds to a redshift, $z\simeq 6.68$.

>>>>

\def\SKASA{\altaffilmark{2}}
\def\SKASAtxt{\altaffiltext{2}{Square Kilometre Array South Africa (SKA SA), Cape Town 7405, South Africa}}

\def\ANU{\altaffilmark{3}}
\def\ANUtxt{\altaffiltext{3}{Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia}}

\def\Haystack{\altaffilmark{4}}
\def\Haystacktxt{\altaffiltext{4}{MIT Haystack Observatory, Westford, MA 01886, USA}}

\def\RRI{\altaffilmark{5}}
\def\RRItxt{\altaffiltext{5}{Raman Research Institute, Bangalore 560080, India}}

\def\Curtin{\altaffilmark{6}}
\def\Curtintxt{\altaffiltext{6}{International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia}}

\def\USydney{\altaffilmark{7}}
\def\USydneytxt{\altaffiltext{7}{Sydney Institute for Astronomy, School of Physics, The University of Sydney, NSW 2006, Australia}}

\def\CAASTRO{\altaffilmark{8}}
\def\CAASTROtxt{\altaffiltext{8}{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}}

\def\MIT{\altaffilmark{9}}
\def\MITtxt{\altaffiltext{9}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}}

\def\CfA{\altaffilmark{10}}
\def\CfAtxt{\altaffiltext{10}{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA 02138, USA}}

\def\UW{\altaffilmark{11}}
\def\UWtxt{\altaffiltext{11}{Department of Physics, University of Washington, Seattle, WA 98195, USA}}

@@

%% Author list
\author{
%% Lead Authors
Nithyanandan~Thyagarajan\ASU,
Daniel~C.~Jacobs\ASU,
Judd~D.~Bowman\ASU,
%% Supporting authors
% A.~P.~Beardsley\UW,
% J.~Pober\UW,
% I.~S.~Sullivan\UW,
%% Builders' list
G.~Bernardi\SKASA,
F.~Briggs\ANU,
R.~J.~Cappallo\Haystack, 
B.~E.~Corey\Haystack, 
A.~A.~Deshpande\RRI, 
D.~Emrich\Curtin,
B.~M.~Gaensler\USydney$^,$\CAASTRO, 
R.~Goeke\MIT,
L.~J.~Greenhill\CfA,
B.~J.~Hazelton\UW, 
M.~Johnston-Hollitt\Victoria,
D.~L.~Kaplan\UWisc, 
J.~C.~Kasper\UMichigan$^,$\CfA, 
E.~Kratzenberg\Haystack, 
C.~J.~Lonsdale\Haystack, 
M.~J.~Lynch\Curtin, 
S.~R.~McWhirter\Haystack,
D.~A.~Mitchell\CASS$^,$\CAASTRO, 
M.~F.~Morales\UW, 
E.~Morgan\MIT, 
D.~Oberoi\Tata, 
S.~M.~Ord\Curtin$^,$\CAASTRO,
T.~Prabu\RRI, 
A.~E.~E.~Rogers\Haystack, 
A.~Roshi\NRAO, 
N.~Udaya~Shankar\RRI, 
K.~S.~Srivani\RRI, 
R.~Subrahmanyan\RRI$^,$\CAASTRO, 
S.~J.~Tingay\Curtin$^,$\CAASTRO, 
M.~Waterson\Curtin$^,$\ANU,
R.~B.~Wayth\Curtin$^,$\CAASTRO, 
R.~L.~Webster\UMelbourne$^,$\CAASTRO, 
A.~R.~Whitney\Haystack, 
A.~Williams\Curtin, 
C.~L.~Williams\MIT
}

@@

Detection of 21~cm H{\sc i} emission from the Epoch of Reionization (EoR), at redshifts $z>6$, is limited primarily by foreground emission.  We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function.  Using the delay spectrum technique, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array (MWA) observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings. For signals entering through the primary field of view, compact objects dominate the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space.  Based on these results, we propose that selective down--weighting of baselines based on length, direction, and time will mitigate a large fraction of foreground contamination in redshifted 21~cm power spectrum analysis.

@@

After the CMB epoch, the period in the history of the universe often referred to as the {\it dark ages}, followed the epoch of reionization (EoR). This was a period of non-linear growth of density perturbations and astrophysical evolution. Studying the epoch of reionization holds the key to understanding this evolution. Direct observation of redshifted 21~cm spin flip transition of H{\sc i} has been identified as a useful probe of the EoR \citep{sun72,sco90,mad97,toz00,ili02} and has recently emerged as a very promising experiment to fill the gaps in our understanding of the universe's history.

@@

A key challenge in the statistical detection of power spectrum of the redshifted H{\sc i} signal arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12,dil13}. Considerable effort has been made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum \citep{thy13,pob13,mor12,tro12,liu11,dat10,bow09,liu09}. A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide-field response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes lost by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components most likely to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.

Foregrounds contaminate the EoR H{\sc i} power spectrum by either having intrinsic spectral variability, high instrumental chromaticity, or through approximations in power spectrum analysis. Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity. To decouple our findings from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using the individual baseline--based approach of \citet{pob13,par14}. This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.

Matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the signature of the power spectrum of each baseline. We report two important findings: the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a larger role than previously thought. % We find that the foregrounds most likely to obfuscate the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon. We find that diffuse Galactic emission plays a larger role then previously thought. 
We quantify this contamination by separating the simulations into components based on type of foreground emission. Using these we arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

@@

Interferometer array data known as {\it visibilities}, $V_f(\overline{\mathbf{u}},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\overline{\mathbf{x}}$ and then Fourier transformed along time axis to obtain a spectrum along frequency ($f$) axis. Here, $\overline{\mathbf{u}}\equiv \overline{\mathbf{x}}f/c \equiv \overline{\mathbf{x}}/\lambda$, $c$ is the speed of light and $\lambda$ is the wavelength. $\overline{\mathbf{u}}$ is related to transverse spatial frequency modes ($\overline{\mathbf{k}}_\perp$) of brightness distribution of the sky as $\overline{\mathbf{k}}_\perp \equiv 2\pi\overline{\mathbf{u}}/D(z)$, where $D(z)$ is the transverse comoving distance at a redshift $z$. If $I(\overline{\mathbf{l}},f)$ is the emission at different frequencies on the sky as a function of direction unit vector ($\overline{\mathbf{l}}$) specified as direction cosines, $V_f(\overline{\mathbf{u}},f)$ represents the Fourier decomposition of $I(\overline{\mathbf{l}},f)$ attenuated by the antenna's angular power--pattern $A(\overline{\mathbf{l}},f)$, into transverse spatial frequency modes. Taking into account the instrumental bandpass weights $W_f(f)$, this relation can be written as:
\begin{align}\label{eqn:obsvis}
  V_f(\overline{\mathbf{u}},f) &= \iint\limits_\textrm{sky} A(\overline{\mathbf{l}},f)\,I(\overline{\mathbf{l}},f)\,W_f(f)\,e^{-i2\pi \overline{\mathbf{u}}\cdot\overline{\mathbf{l}}}\,\dif^2 \overline{\mathbf{l}}.
\end{align}

We define the {\it delay spectrum} $V_\eta(\overline{\mathbf{u}},\eta)$ to be the inverse Fourier transform of $V_f(\overline{\mathbf{u}},f)$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  V_\eta(\overline{\mathbf{u}},\eta) &= \int V_f(\overline{\mathbf{u}},f)\,W'_f(f)\,e^{i2\pi\eta f}\,\dif f,
\end{align}
where, $W'_f(f)$ is a frequency window weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}. $\eta$ represents the signal delay between antenna pairs:
\begin{equation}\label{eqn:delay}
  \eta = \frac{\overline{\mathbf{x}}\cdot\overline{\mathbf{l}}}{c} = \frac{\overline{\mathbf{u}}\cdot\overline{\mathbf{l}}}{f}.
\end{equation}
Note that as the visibility is a fully complex quantity, its Fourier transform is fully Nyquist sampled over positive and negative delays. Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna spacing vector. In this paper, $\overline{\mathbf{u}}$ and $\overline{\mathbf{x}}$ are assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site. Hence, a perfectly Eastward oriented antenna spacing will observe an object in the Eastern sky at positive delay, and conversely for an object in the Western sky. Similarly, an object in the Northern sky will appear at a positive delay at an antenna spacing oriented Northward.

Since we are studying a redshifted H{\sc i} spectral line, $f$ is a measure of cosmological distance along the line of sight. Consequently, $\eta$ is a measure of line of sight spatial frequency mode given by $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. We use $H_0=100\,h$~km~s$^{-1}$~Mpc$^{-1}$ and assume a flat $\Lambda$CDM universe with $\Omega_\textrm{M}=0.27$ and $\Omega_\Lambda=0.73$.

Thus, $|V_\eta(\overline{\mathbf{u}},\eta)|^2$ is directly related to the spatial power spectrum of EoR H{\sc i} distribution, $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ \citep{mor04}. However, foreground objects also emit in radio frequencies. Due to their spatial and spectral (chromatic) structure, the measured $V_\eta(\overline{\mathbf{u}},\eta)$ is contaminated by foreground radio objects \citep{thy13,tro12,mor12,bow09}. These contaminations have to be characterized precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity.

In summary, $V_\eta(\overline{\mathbf{u}},\eta)$ is closely related to visibilities, $V_f(\overline{\mathbf{u}},f)$, the basic data blocks measured by each antenna pair, as seen from equation~\ref{eqn:delay-transform}. It captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is directly related to the sought power $|V(\overline{\mathbf{k}}_\perp,k_\parallel)|^2$ in $k$--modes containing critical information about spatial scales. For our analysis, we use an equivalent quantity, $V_\eta(\overline{\mathbf{x}},\eta) \equiv V_\eta(\lambda\overline{\mathbf{u}},\eta)$. For purposes of illustration, we often use a simplified quantity $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$.

Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper. $|\overline{\mathbf{x}}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis. $\eta$ and $k_\parallel$, denoting spatial scales along line of sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge}, the boundaries of which are determined by antenna spacing and the light travel time across it. These boundaries are called horizon delay limits and are shown by solid lines. Spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\eta$--axis. This effect has been quantified in \citet{thy13}. The width of this narrow strip is inversely proportional to operating bandwidth. Region of Fourier space excluding the {\it foreground wedge} and the narrow strips is usually called the {\it EoR window}. These are shown in light and medium shades. However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape giving rise to periodic gratings thereby resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s. The underlying cause of this effect will be discussed later in this paper.

% Hence, we have identified a region of high significance (shown in medium shade) that forms a subset of the {\it EoR window} bounded by $|\eta|\le 0.78\,\mu$s where contamination from foregrounds is expected to be less and yet the EoR H{\sc i} power spectrum signal is reasonably strong. Hereafter, we refer to this region as the {\it EoR window} in the rest of this paper.

In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\overline{\mathbf{k}}|$. Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays. Thus the region of interest for EoR studies is slightly beyond the horizon delay limits (dashed lines) and inside the first grating on either side ($|\eta| \lesssim 0.78\,\mu$s, dot--dashed lines) caused by the instrument's spectral transfer function. We call this region with medium shade as the {\it EoR window} in this paper. The MWA EoR collaboration actively uses this as a test region in Fourier space for diagnostics on data quality.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
\caption{Illustration of Fourier space in which delay (and power) spectra of H{\sc i} signals from the EoR are presented. The $x$--axis denoted by $|\overline{\mathbf{x}}|$ (antenna spacing in m) or $k_\perp$ (transverse wavenumber in units of $h$~Mpc$^{-1}$) are equivalent and represent transverse spatial scales. Similarly, the $y$--axis denoted by $\eta$ (delay in s or $\mu$s) or equivalently, $k_\parallel$ (line--of--sight wavenumber in units of $h$~Mpc$^{-1}$) represents spatial scales along the line of sight. The dark shaded region is often referred to as {\it foreground wedge} where smooth spectrum foregrounds reside. The boundaries of the {\it foreground wedge} (solid lines) are given by the light travel time for corresponding antenna spacings and are referred to as horizon delay limits. In principle, region outside the wedge is expected to be relatively free of foreground contamination and is called the {\it EoR window} (light and medium shades). In practice, this contamination is also high in the narrow unshaded white regions (bound by solid and dashed lines) due to convolution by instrumental transfer function. The width of this narrow region is determined by inverse of operating bandwidth. Due to an undesirable spectral response of the instrument used in this study, namely the MWA, we identify a narrower region of high significance to EoR studies (medium shade) and hereafter refer to it as the {\it EoR window} in this paper. \label{fig:fourier-space}}
\end{figure}

The terms antenna spacing and baseline are synonymous in radio interferometry. The latter term is also used broadly to refer to the antenna pair system in a radio interferometer array. We use these terms interchangeably in this paper.

For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$. Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.

@@

The MWA was completed in 2012, and after commissioning, began its EoR observing program in 2013. In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30~MHz band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles each in the shape of a bow--tie. This yields a primary field of view $\gtrsim$~20\arcdeg wide and multiple secondary lobes for the tile. The tile arrangement is a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius with the longest baselines at 3~km \citep{bea12}. Here we focus on antenna spacings $|\overline{\mathbf{x}}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz. The center frequency 185~MHz corresponds to a redshift, $z\simeq 6.68$.
